Line of Sight Lensing Mass Reconstructions in the Millennium Simulation

### 2016-06-05 doc/thesis_swe.tex
Thesis update. All captions and citations added. #98

\makeatletter
\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark} % Used for footnotes
\def\p@subsection{} % Fixes section references
\makeatother

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Millennium Siulation structure
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/thesis/ms_structure.jpg}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{\cite{ms_figures}}
    \label{fig:ms_structure}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% FIGURE: Millennium Simulation Distribution plots (cosmic web, galaxy clusters with/without dark matter)
\begin{figure*}
    \centering
    \begin{subfigure}[H]{0.375\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ms_galaxies.jpg}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{ms_galaxies}
    \end{subfigure}
    ~
    %% Einstein Ring Figure
    \begin{subfigure}[H]{0.375\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ms_halos.jpg}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{ms_halos}
    \end{subfigure}
    \caption{\cite{ms_figures}}
    \label{ms_compare}
\end{figure*}

@@

Cosmological simulations like the Millennium Simulation (MS) demonstrate that the underlying matter distribution of our universe is likely breathtakingly complex; rather than a largely disordered collection of galaxies and interstellar gas, matter appears to condense into clusters and superclusters of galaxies threaded together by large filaments pocketed with under-dense regions called voids, as shown in Figure \ref{fig:ms_structure} \cite{ms_figures}. While the exact relation between the distribution of galaxies and dark matter is not known, simulations predict that early galaxies formed in the large gravitational potentials of over-dense regions of dark matter \cite{halo_formation}. This suggests that, while we cannot map dark matter directly, galaxies should trace out at least some of the underlying dark matter structures \cite{abundance_matching}. A comparrison of the predicted galaxy and dark matter distribution from a small region of the MS is shown in Figure \ref{ms_compare}.

@@

 The analytical form of the convergence and shear contribution from a BMO-profile is quite messy, but is derived and explained in detail in \ref{bmo}. Calculating the lensing contribution of the smooth-component correction is much simpler; the projected surface mass density at at given redshift is found by integrating $\rho_s(z)$ over a sufficiently small $dz$ interval, and then the mass sheet's convergence is found using Equation \eqref{convergence}. There is no shear contribution as a uniform sheet of mass causes only magnification due to the mass sheet degeneracy \cite{mass_degeneracy}. However, Equation \eqref{lensing} shows that the lensed ellipticities depend on the \textit{reduced} shear, which does depend on the additional convergence terms resulting from the smooth-component correction.

@@

To infer foreground dark matter mass structures using weak gravitational lensing, first a model of the relationship between foreground galaxies and dark matter must be established and robustly tested to see if, statistically, it makes the same lensing predictions of background sources as the true underlying dark matter structure. To do this, I built upon the publicly available \texttt{Pangloss}\footnote{\label{note1}https://github.com/drphilmarshall/Pangloss} framework used in Collett et al. (\cite{collett}) to reconstruct all the
mass along the lines of sight of generated background galaxies in the MS using a halo model with smooth-component correction. The lensing contribution of each halo is calculated, and the total lensing of the background galaxy is the sum of each halo contribution. The implementation of this process is detailed in the following sections.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Generating Background Galaxies}
With the (0,0,0,0) catalog of MS foreground galaxies chosen, a set of 1440 background galaxies over a field of 144 arcmin$^2$ was generated, with density 10 galaxies per arcmin$^2$. The intrinsic orientation of each galaxy was sampled from a uniform distribution as, without lensing, there should be no preferred orientation due to the assumption of an isotropic universe. The magnitude of the galaxy ellipticities was sampled from a normal distribution with a standard deviation of 0.2, but any ellipticities with magnitude greater than one were re-sampled. Random ellipticity noise was sampled from a normal distribution with a standard deviation of 0.1 and added to the intrinsic ellipticities. Finally, each ellipticity was multiplied by $M=0.9$ to account for a multiplicative bias of 10\%.

@@

Not all foreground halos are equally relevant to the lensing calculation. To add a significant contribution to the combined shear and convergence, a halo must be massive, close to the line of sight, or preferably both. McCully et al. in \ref{mccully} derived that the correct metric for determining the relevance of a given foreground halo to the overall lensing calculation to be given by
\begin{equation}\label{relevant}
Rel(M_h,R)\propto\frac{M_h}{R^3}
\end{equation}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Relevance Distribution of all Halos
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs-swe/thesis/relevance_distribution.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:rel_dist}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Smooth-Component Correction Diagram
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/thesis/smooth_correction.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:scc}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

The volume and projected surface mass densities of each redshift slice are shown in Figures \ref{fig:volume_densities} and \ref{fig:surface_densities} respectively. In each plot, the halo density at each redshift slice is plotted in green, the smooth-component density in blue, and universe mean density in black, and a histogram of the galaxies per bin is shown in the background. In most redshift bins, the halo density is one to two orders of magnitude lower than than the universe mean density leading to a smooth-component density very close to that of the mean density. However, all mass densities are far below the critial density, plotted in red, and so well within the weak-lensing regime. The convergence distribution of each halo and 100 smooth-components is plotted in Figure \ref{fig:kappa_dist}.
% This seems to be in opposition to Pace et al. in \cite{pace}, which found that halos constituted ??\% of simulation mass in ??.
%%------------------------------------------------------------------------------------------------------------------------------------------
%% Surface and Volume densities for smooth-component correction redshift slices
\begin{figure*}
    \centering
    % Volume
    \begin{subfigure}[H]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/vol_densities.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{fig:volume_densities}
    \end{subfigure}
    ~
    % Surface
    \begin{subfigure}[H]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/surf_densities.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{fig:surface_densities}
    \end{subfigure}
    \caption{}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Kappa Contributions of S.C.C.
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/thesis/kappa_distribution.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:kappa_dist}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

Using the physical separation distance and halo mass, the shear and convergence contribution of a single foreground halo is calculated using methods described in Section \ref{grav_lensing} (see Wright and Brainerd \cite{lensing_calc} for far more detail). The total convergence and shear along the LOS of the lightcone is simply the sum of the convergence and shear contributions of each halo contained in the lightcone, as well as the additional convergence terms from the smooth-component correction. The \texttt{Pangloss} predicted lensed source ellipticity is then calculated using Equation \eqref{lensing}.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Pangloss Lightcone Data Plots
\begin{figure*}
    \centering
    \includegraphics[width=0.75\textwidth]{figs-swe/thesis/lightcone_plots.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:lightcone}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

Figure \ref{fig:lightcone} displays different views an example lightcone where the foreground halos are plotted as circles whose size are proportional to their galaxy brightness (top left), angular size (top center), convergence contributions (top right), and halo mass (bottom). Figure \ref{fig:subplots} shows a few examples of mass structures in the foreground catalog along with the ray-traced convergence (plotted in inverse greyscale) and shear maps (plotted as red ellipticity sticks) from \cite{ray_tracing}.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
% Ellipticity-Ellipticity Correlation
\begin{figure}
    \centering
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/gg_corr.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        %The ellipticity-ellipticity correlation function for ?? sources at separation distances from 0.1 to 2 arcminutes and a lightcone radius of $R=2$ arcminutes. While the correlation of the intrinsic ellipticities before lensing (blue) shows no correlation, the correlation of the ray-traced ellipticities (green) and halo model predicted ellipticities (purple) both show significant positive correlation on scales up to 1 arcminute. The halo model predicts a small negative correlation above this scale, while the ray-tracing shows positive correlation on all scales.}
        %The correlation of the intrinsic ellipticities before lensing is plotted in blue, and shows no correlation. The correlation of the true ray-traced ellipticities is plotted in green and shows significant positive correlation on all scales. The halo predicted correlation, plotted in purple, shows significant positive correlation until a separation distance of 0.8 arcminutes.
        \label{fig:gg_corr}
    \end{subfigure}
    ~~
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/gg_progression.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        %The ellipticity-ellipticity correlation functions for a series of halo model frameworks with increasing lightcone radii from $R=1$ to $R=8$ arcminutes, all plotted in purple with various line styles. These are compared to the same ray-traced correlation from Figure \ref{gg_corr}.}
        \label{fig:gg_corr_series}
    \end{subfigure}
    \caption{}
\end{figure}
%------------------------------------------------------------------------------------------------------------------------------------------

The first test of the halo model was with the $\varepsilon$-$\varepsilon$ correlation function with a lightcone radius of 6 arcminutes which is given in Figure \ref{fig:gg_corr}. The statistic measured the average correlation between pairs of ellipticities at separation distances between 0.1 and 2 arcminutes. The cross-component of the correlation function shows no significant deviation from zero as expected, as gravitational lensing only shears galaxies tangentially. For the $\xi_+$ component, the ray-traced (plotted in green) and halo model (plotted in purple) are in relative agreement from seperation scales of 0.1 to 0.2 arcminutes, but there is a systematic underprediction of correlation by the halo model on scales larger than 0.2 arcminutes by an average of 31\%. This result is consistent with Pace et al. in \cite{pace} which found that a purely halo model underpredicted the lensing correlation on average by 40\%. This may be indicative that not enough halo mass is being used in the lensing prediction, the halo model does not adequately address large-scale mass structures, or that there is significant dark matter mass not correlated with galaxies.  These issues are discussed further in Section \ref{issues}.

@@

The catalog of lensed ellipticities was also analyzed using the galaxy-mass correlation function with separation distances from 0.1 to 2 arcminutes and a lightcone radius of 6 arcminutes. The result is shown in Figure \ref{fig:ng_corr}. The lensed ellipticities from the ray-tracing and halo model both show positive correlation on all calculated scales and largely agree on scales larger than 0.3 arcminutes except for a slight overprediction by the halo model near 0.45 arcminutes. However, a new feature is the large overprediction of correlation on separation scales smaller than 0.2 arcminutes. This is likely because background sources this close to massive foreground halos are often in the strong lensing regime which is not currently accounted for by the \texttt{Pangloss} framework.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Galaxy-mass correlation function
\begin{figure}
    \centering
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ng_corr.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        %The galaxy-mass correlation function of 3,240 sources at separation distances from 0.1 to 2 arcminutes and a lightcone radius of $R=2$ arcminutes. There is no significant correlation of the intrinsic ellipticities (blue) as expected, and both the ray-traced correlation (green) and halo model correlation (purple) are positive on all scales. The green and purple curves are in agreement from 0.2 to 0.7 arcminutes, but have significant discrepancies on scales above and below this range.}
        \label{fig:ng_corr}
    \end{subfigure}
    ~~
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ng_progression.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{fig:ng_corr_series}
    \end{subfigure}
    \caption{}
\end{figure}
%
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

Of intrinsic interest is the relative number of relevant halos needed in each lightcone to accurately predict the halo correlation functions in Figures \ref{fig:gg_corr} and \ref{fig:ng_corr}, as a low percentage would allow much faster lensing predictions. Figures \ref{fig:rel_gg_compare} and \ref{fig:rel_ng_compare} display an example $\varepsilon$-$\varepsilon$ and galaxy-mass correlation function comparrison of the halo model using all halos (purple) and halo model using halos above the relevance limit of ${10^{-5} M_{\odot}/\text{Mpc}^3}$, as well as the percent error of each predicted value between the two implementations. This relevance limit resulted in $??\pm??$ relevant halos per lightcone, and resulted in a normalized root-mean-square error (NRMSE) of ??. Figure \ref{fig:halo_fraction} plots the $\varepsilon$-$\varepsilon$ and galaxy-mass correlation function NRMSE as a function of relevance limits and mean relevant halos. These results suggest that either correlation function prediction can be approximated on average with an NRMSE below 10\% using around the 70 most relevant halos, or an NRMSE of below 5\% using around the 300 most relevant halos. Calibration curves such as this can allow the best-case correlation function to be predicted within a desired accuracy at only a fraction of the full lensing calculation which will be especially helpful in future analyses with much larger fields of view.

>>>>

\makeatletter
\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark} % Used for footnotes
\def\p@subsection{} % Fixes section references
\renewcommand\@make@capt@title[2]{%
     \@ifx@empty\float@link{\@firstofone}{\expandafter\href\expandafter{\float@link}}%
      {\textbf{#1}}\@caption@fignum@sep#2\quad}% % Fixes caption justification
\makeatother

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Millennium Siulation structure
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/thesis/ms_structure.jpg}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{A visualization of the large-scale dark matter distribution in the Millennium Simulation. The largest bright spots correspond to superclusters of galaxies which are connected by thread-like structures of matter called filaments. Taken from \cite{ms_figures}.}
    \label{fig:ms_structure}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% FIGURE: Millennium Simulation Distribution plots (cosmic web, galaxy clusters with/without dark matter)
\begin{figure*}
    \centering
    \begin{subfigure}[H]{0.375\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ms_galaxies.jpg}
        %\captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{Galaxy distribution.}
        \label{ms_galaxies}
    \end{subfigure}
    ~
    %% Einstein Ring Figure
    \begin{subfigure}[H]{0.375\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ms_halos.jpg}
        %\captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{Halo distribution.}
        \label{ms_halos}
    \end{subfigure}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{A comparison of the galaxy and dark matter distribution for a single cluster in the Millennium Simulation. The galaxies tend to be located in the centers of clumps of dark matter called `halos', which suggests that galaxies at least partially trace out the underlying dark matter distribution. Taken from \cite{ms_figures}.}
    \label{ms_compare}
\end{figure*}

@@

Cosmological simulations like the Millennium Simulation demonstrate that the underlying matter distribution of our universe is likely breathtakingly complex; rather than a largely disordered collection of galaxies and interstellar gas, matter appears to condense into clusters and superclusters of galaxies threaded together by large filaments pocketed with under-dense regions called voids, as shown in Figure \ref{fig:ms_structure} \cite{ms_figures}. While the exact relation between the distribution of galaxies and dark matter is not known, simulations predict that early galaxies formed in the large gravitational potentials of over-dense regions of dark matter \cite{halo_formation}. This suggests that, while we cannot map dark matter directly, galaxies should trace out at least some of the underlying dark matter structures \cite{abundance_matching}. A comparison of the predicted galaxy and dark matter distribution from a small region of the simulation is shown in Figure \ref{ms_compare}.

@@

 The analytical form of the convergence and shear contribution from a BMO-profile is quite messy, but is derived and explained in detail in \ref{nfw_bmo}. Calculating the lensing contribution of the smooth-component correction is much simpler; the projected surface mass density at at given redshift is found by integrating $\rho_s(z)$ over a sufficiently small $dz$ interval, and then the mass sheet's convergence is found using Equation \eqref{convergence}. There is no shear contribution as a uniform sheet of mass causes only magnification due to the mass sheet degeneracy \cite{mass_degeneracy}. However, Equation \eqref{lensing} shows that the lensed ellipticities depend on the \textit{reduced} shear, which does depend on the additional convergence terms resulting from the smooth-component correction.

@@

To infer foreground dark matter mass structures using weak gravitational lensing, first a model of the relationship between foreground galaxies and dark matter must be established and robustly tested to see if, statistically, it makes the same lensing predictions of background sources as the true underlying dark matter structure. To do this, I built upon the publicly available \texttt{Pangloss}\footnote{\label{note1}https://github.com/drphilmarshall/Pangloss} framework used in Collett et al. (\cite{collett}) to reconstruct all the
mass along the lines of sight of generated background galaxies in the Millennium Simulation using a halo model with smooth-component correction. The lensing contribution of each halo is calculated, and the total lensing of the background galaxy is the sum of each halo contribution. The implementation of this process is detailed in the following sections.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Generating Background Galaxies}
With the (0,0,0,0) catalog of Millennium Simulation foreground galaxies chosen, a set of 1440 background galaxies over a field of 144 arcmin$^2$ was generated, with density 10 galaxies per arcmin$^2$. The intrinsic orientation of each galaxy was sampled from a uniform distribution as, without lensing, there should be no preferred orientation due to the assumption of an isotropic universe. The magnitude of the galaxy ellipticities was sampled from a normal distribution with a standard deviation of 0.2, but any ellipticities with magnitude greater than one were re-sampled. Random ellipticity noise was sampled from a normal distribution with a standard deviation of 0.1 and added to the intrinsic ellipticities. Finally, each ellipticity was multiplied by $M=0.9$ to account for a multiplicative bias of 10\%.

@@

Not all foreground halos are equally relevant to the lensing calculation. To add a significant contribution to the combined shear and convergence, a halo must be massive, close to the line of sight, or preferably both. McCully et al. in \cite{mccully} derived that the correct metric for determining the relevance of a given foreground halo to the overall lensing calculation to be given by
\begin{equation}\label{relevant}
Rel(M_h,R)\propto\frac{M_h}{R^3}
\end{equation}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Relevance Distribution of all Halos
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs-swe/thesis/relevance_distribution.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{The mean number of relevant galaxies in a single realization of 1440 source galaxies determined by Equation \eqref{relevant_thresh} with a threshold mass and distance of ${M_T=10^{12}M_\odot}$ and ${R_T=10}$ kpc respectively. The errorbars correspond to the standard deviation of relevant halos accross all lightcones. As expected, most halos are either too far away from the line-of-sight or not massive enough to make a significant contribution, or both. The number of halos falls logarithmically with the relevance limit and reaches a mean of one halo at a relevance of ${10^{-2}M_\odot\,\text{Mpc}^3}$.}
    \label{fig:rel_dist}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Smooth-Component Correction Diagram
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/thesis/smooth_correction.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{A diagram of the smooth-component correction implementation in \texttt{Pangloss}. The halo density at each redshift slice $z$ was calculated by summing the masses of all foreground halos within $\Delta z$ and dividing by the proper volume of the cone slice using Equation \eqref{pvolume_element} and solid angle $d\Omega$.}
    \label{fig:scc}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

The volume and projected surface mass densities of each redshift slice are shown in Figures \ref{fig:volume_densities} and \ref{fig:surface_densities} respectively. In each plot, the halo density at each redshift slice is plotted in green, the smooth-component density in blue, and universe mean density in black, and a histogram of the galaxies per bin is shown in the background. In most redshift bins, the halo density is one to two orders of magnitude lower than than the universe mean density leading to a smooth-component density very close to that of the mean density. However, all mass densities are far below the critial density, plotted in red, and so well within the weak-lensing regime. The convergence distribution of each halo and 100 smooth-components is plotted in Figure \ref{fig:kappa_dist}.
% This seems to be in opposition to Pace et al. in \cite{pace}, which found that halos constituted ??\% of simulation mass in ??.
%%------------------------------------------------------------------------------------------------------------------------------------------
%% Surface and Volume densities for smooth-component correction redshift slices
\begin{figure*}
    \centering
    % Volume
    \begin{subfigure}[H]{0.48\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/vol_densities.png}
        % \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{Smooth-component volume densities.}
        \label{fig:volume_densities}
    \end{subfigure}
    ~
    % Surface
    \begin{subfigure}[H]{0.48\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/surf_densities.png}
        % \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{Smooth-component projected surface densities.}
        \label{fig:surface_densities}
    \end{subfigure}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{The surface and volume densities for each of the 100 smooth-component correction and halo densities in a lightcone with radius $R_L=6$ arcminutes. The green line is the halo density of each slice and the blue line is the calculated smooth-component density such that the two densities sum to the universe mean density plotted in black. The number of halos in each redshift bin is plotted as a histogram in the background. The red line in plot (b) is the ciritical surface density.}
    \label{densities}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Kappa Contributions of S.C.C.
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/thesis/kappa_distribution.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{The convergence distribution of the 4838 individual halos, plotted in green, and the 100 smooth-component slices, plotted in blue, contained in the same lightcone as Figure \ref{densities}. The mass sheets have a higher convergence contribution than all but the most relevant halos in the lightcone. The curve of smooth component $\kappa$'s indicates how foreground mass at the center of the light's path has the highest lensing efficiency.}
    \label{fig:kappa_dist}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

Using the physical separation distance and halo mass, the shear and convergence contribution of a single foreground halo is calculated using methods described in Section \ref{grav_lensing} (see \cite{lensing_calc} and \cite{nfw_bmo} for more details). The total convergence and shear along the LOS of the lightcone is simply the sum of the convergence and shear contributions of each halo contained in the lightcone, as well as the additional convergence terms from the smooth-component correction. The \texttt{Pangloss} predicted lensed source ellipticity is then calculated using Equation \eqref{lensing}.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Pangloss Lightcone Data Plots
\begin{figure*}
    \centering
    \includegraphics[width=0.75\textwidth]{figs-swe/thesis/lightcone_plots.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{Four different views of a lightcone centered on the line-of-sight of a particular background source.  The top left figure plots the positions of the halos with the size proportional to their galaxy brightness in the $i$-band, the center figure plots the angular size of foreground halos with red and blue regions representing the NFW scale radius and virial radius of each halo respectively, the top right figure plots the $\kappa$ contributions of each halo, and the bottom plots the positions of each halo along the redshift axis with size proportional to halo mass.}
    \label{fig:lightcone}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

Figure \ref{fig:lightcone} displays different views an example lightcone where the foreground halos are plotted as circles whose size are proportional to their galaxy brightness (top left), angular size (top center), \texttt{Pangloss} predicted convergence contributions (top right), and halo mass (bottom). Comparing the center and top right figures demonstrates the importance of proximity to the line-of-sight for significant $\kappa$ contributions. Figure \ref{fig:subplots} shows a few examples of mass structures in the foreground catalog along with the ray-traced convergence (plotted in inverse greyscale) and shear maps (plotted as red ellipticity sticks) from \cite{ray_tracing}.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
% Ellipticity-Ellipticity Correlation
\begin{figure}
    \centering
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/gg_corr.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The $\varepsilon$-$\varepsilon$ correlation function for 1440 sources at separation distances from 0.1 to 2 arcminutes using lightcones with radii of $R_L=6$ arcminutes, where green is the `true' ray-traced correlation and purple is the halo model predicted correlation. While the cross-component $\xi_\times(\Delta\theta)$ for both is consistent with zero accross all scales as expected, the halo model systematically underpredicts the correlation on scales larger than 0.2 arcminutes.}
        \label{fig:gg_corr}
    \end{subfigure}
    ~~
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/gg_progression.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The $\varepsilon$-$\varepsilon$ correlation functions of 1440 sources for a series of halo model frameworks with increasing lightcone radii from $R_L=1$ to $R_L=8$ arcminutes, all plotted in purple with various line styles. These are compared to the same ray-traced correlation from Figure \ref{fig:gg_corr}. The predicted correlation seems to converge to the ray-traced value with larger $R_L$ as more halo contributions are accounted for.}
        \label{fig:gg_corr_series}
    \end{subfigure}
    \caption{}
\end{figure}
%------------------------------------------------------------------------------------------------------------------------------------------

The first test of the halo model was with the $\varepsilon$-$\varepsilon$ correlation function with a lightcone radius of 6 arcminutes and all halos which is given in Figure \ref{fig:gg_corr}. The statistic measured the average correlation between pairs of ellipticities at separation distances between 0.1 and 2 arcminutes. The cross-component of the correlation function shows no significant deviation from zero as expected, as gravitational lensing only shears galaxies tangentially. For the $\xi_+$ component, the ray-traced (plotted in green) and halo model (plotted in purple) are in relative agreement from seperation scales of 0.1 to 0.2 arcminutes, but there is a systematic underprediction of correlation by the halo model on scales larger than 0.2 arcminutes by an average of 31\%. This result is consistent with Pace et al. in \cite{pace} which found that a purely halo model underpredicted the lensing correlation on average by 40\%. This may be indicative that not enough halo mass is being used in the lensing prediction, the halo model does not adequately address large-scale mass structures, or that there is significant dark matter mass not correlated with galaxies.  These issues are discussed further in Section \ref{issues}.

@@

The catalog of lensed ellipticities was also analyzed using the galaxy-mass correlation function with separation distances from 0.1 to 2 arcminutes and a lightcone radius of 6 arcminutes and all halo. The result is shown in Figure \ref{fig:ng_corr}. The lensed ellipticities from the ray-tracing and halo model both show positive correlation on all calculated scales and largely agree on scales larger than 0.3 arcminutes except for a slight overprediction by the halo model near 0.45 arcminutes. However, a new feature is the large overprediction of correlation on separation scales smaller than 0.2 arcminutes. This is likely because background sources this close to massive foreground halos are often in the strong lensing regime which is not currently accounted for by the \texttt{Pangloss} framework.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Galaxy-mass correlation function
\begin{figure}
    \centering
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ng_corr.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The galaxy-mass correlation function of 1440 sources at separation distances from 0.1 to 2 arcminutes and a lightcone radius of $R_L=6$ arcminutes. The halo model predicts the statistic well on scales above 0.3 arcminutes, but systematically overpredicts correlation on scales below 0.2 arcminutes. This is likely due to strong lensing effects which are not currently accounted for in \texttt{Pangloss}.}
        \label{fig:ng_corr}
    \end{subfigure}
    ~~
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ng_progression.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The galaxy-mass correlation functions of 1440 sources for a series of halo model frameworks with increasing lightcone radii from $R_L=1$ to $R_L=8$ arcminutes, all plotted in purple with various line styles. These are compared to the same ray-traced correlation from Figure \ref{fig:ng_corr}. The predicted correlation seems to converge to the ray-traced value with larger $R_L$ as more halo contributions are accounted for.}
        \label{fig:ng_corr_series}
    \end{subfigure}
    \caption{}
\end{figure}
%
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

Of intrinsic interest is the relative number of relevant halos needed in each lightcone to accurately predict the halo correlation functions in Figures \ref{fig:gg_corr} and \ref{fig:ng_corr} as a low percentage would allow much faster lensing predictions. Figures \ref{fig:rel_gg_compare} and \ref{fig:rel_ng_compare} display an example $\varepsilon$-$\varepsilon$ and galaxy-mass correlation function comparison of the halo model using all halos (plotted in purple) and halo model using halos above the relevance limit of ${10^{-5} M_{\odot}/\text{Mpc}^3}$ (plotted in blue) with $R_L=6$ arcminutes, as well as the percent error of each predicted value between the two implementations. This relevance limit corresponded to $66\pm10$ relevant halos per lightcone and resulted in a normalized root-mean-square error (NRMSE) of 0.164 and 0.082 for the two statistics. Figure \ref{fig:halo_fraction} plots the $\varepsilon$-$\varepsilon$ and galaxy-mass correlation function NRMSE as a function of relevance limits and mean relevant halos, still using a lightcone radius of 6 arcminutes. These results suggest that either correlation function prediction can be approximated on average with an NRMSE below 10\% using around the 70 most relevant halos, or an NRMSE of below 5\% using around the 300 most relevant halos. Calibration curves such as this can allow the best-case correlation function to be predicted within a desired accuracy at only a fraction of the full lensing calculation which will be especially helpful in future analyses with much larger fields of view.

%%------------------------------------------------------------------------------------------------------------------------------------------
% Correlation Function comparison
\begin{figure}
    \centering
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/rel_gg_compare.png}
        % \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{$\varepsilon$-$\varepsilon$ correlation function comparison.}
        \label{fig:rel_gg_compare}
    \end{subfigure}
    ~~
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/rel_ng_compare.png}
        % \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{Galaxy-mass correlation function comparison.}
        \label{fig:rel_ng_compare}
    \end{subfigure}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{Comparing the two correlation functions using all halos (purple) and only the $66\pm10$ most relevant halos resulting from a relevance limit of ${10^{-5} M_{\odot}/\text{Mpc}^3}$ (blue). The percent error of each point as well as the mean error and NRMSE is plotted below each figure.}
    \label{fig:rel_corr}
\end{figure}
